---
title: "Lab 02 - Plastic waste"
author: "Yuxin Xie"
date: "1/28/2025"
output: github_document
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
plastic_waste <- read.csv("data/plastic-waste.csv")
```

## Exercises

### Exercise 1

```{r plastic-waste-continent}
ggplot(data = plastic_waste, aes(x = plastic_waste_per_cap)) +
  geom_histogram(binwidth = 0.2)
```

```{r identify-outlier}
plastic_waste %>%
  filter(plastic_waste_per_cap > 3.5)
```
#### I did not expect this result, likely because this small island nation has high levels of consumption but limited recycling infrastructure. 


```{r 1.1.1}
ggplot(data = plastic_waste, aes(x = plastic_waste_per_cap)) +    facet_wrap(~continent) + geom_histogram(binwidth = 0.2) 
```
# 5 out of 6 continents' plastic waste per capita are within the 0-1 range. North America show higher variation, with Trinidad and Tobago having significantly higher plastic waste per capita.

```{r 1.1.2}
ggplot(
  data = plastic_waste,
  aes(x = plastic_waste_per_cap) ) +
  geom_density()
```

```{r 1.1.3}
ggplot(
  data = plastic_waste,
  mapping = aes(
    x = plastic_waste_per_cap,
    color = continent
  )
) +
  geom_density()
```

```{r 1.1.4}
ggplot(
  data = plastic_waste,
  mapping = aes(
    x = plastic_waste_per_cap,
    color = continent,
    fill = continent
  )
) +
  geom_density()
```

```{r 1.1.5}
ggplot(
  data = plastic_waste,
  mapping = aes(
    x = plastic_waste_per_cap,
    color = continent,
    fill = continent
  )
) +
  geom_density(alpha = 0.7)
```

### Exercise 2

```{r plastic-waste-density}
ggplot(
  data = plastic_waste,
  mapping = aes(
    x = plastic_waste_per_cap,
    color = continent,
    fill = continent
  )
) +
  geom_density(alpha = 0.2)
```
# we defined the color and fill of the curves by mapping aesthetics of the plot but we defined the alpha level as a characteristic of the plotting geom. that is because color and fill vary based on continent data. Each continent is assigned a unique color and fill. the alpha level is uniform across all continents.

```{r another way}
ggplot(
  data = plastic_waste,
  mapping = aes(
    x = continent,
    y = plastic_waste_per_cap
  )
) +
  geom_boxplot()
```

### Exercise 3

```{r plastic-waste-violin}
ggplot(
  data = plastic_waste,
  mapping = aes(
    x = continent,
    y = plastic_waste_per_cap
  )
) +
  geom_violin()
```
### the width of the violin varies, which represents the density. 

### Exercise 4

```{r plastic-waste-mismanaged}
ggplot(data = plastic_waste, aes(x = plastic_waste_per_cap, y = mismanaged_plastic_waste_per_cap)) +
  geom_point()
```
## there seems to be a weak positive relationship between plastic waste per capita and mismanaged plastic waste per capita

```{r colored-plastic-waste-mismanaged-continent}
ggplot(data = plastic_waste, aes(x = plastic_waste_per_cap, y = mismanaged_plastic_waste_per_cap, color = continent)) +
  geom_point()
```
### there is no absolute separation between continents. Africa appears to show a clearer positive pattern, while Europe seems more scattered.

```{r plastic-waste-population-total}
ggplot(data = plastic_waste, aes(x = plastic_waste_per_cap, y = total_pop)) +
  geom_point()
```

```{r plastic-waste-population-coastal}
ggplot(data = plastic_waste, aes(x = plastic_waste_per_cap, y = coastal_pop)) +
  geom_point()
```
###Neither pair of variables shows a strong linear relationship. 

### Exercise 5

#

```{r recreate-viz}

```


